---
title: "Validating Experiments in HRI: Descriptive Results"
subtitle: "Paper-level coding analysis (N = 73)"
date: "`r format(Sys.Date(), '%Y-%m-%d')`"
output:
  html_document:
    toc: true
    toc_depth: 2
    df_print: kable
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

## Data

Input: `HRI_Validation_Coding_Data_Paperlevel.csv` (paper-level clean coding
data). One row per paper. For papers reporting more than one experiment,
every Yes/No indicator was computed with an "at least one experiment/IV/DV =
Yes" rule; see the `Data_Dictionary` tab of
`HRI_Validation_Coding_Data_Paperlevel.xlsx` for the exact rule used for
each column.

This document uses base R plus `knitr::kable` for formatting only - no other
packages are required.

```{r load-data}
df <- read.csv("HRI_Validation_Coding_Data_Paperlevel.csv", stringsAsFactors = FALSE)
n_papers <- nrow(df)
n_papers
```

## Sample and design descriptives

```{r sample-descriptives}
knitr::kable(as.data.frame(table(Technical_System_Paper = df$Technical_System_Paper)),
             caption = "Technical_System_Paper distribution")

cat("Total participants across all papers (Total_Sample_Size summed):",
    sum(df$Total_Sample_Size, na.rm = TRUE), "\n")
summary(df$Total_Sample_Size)
```

`Experiment_Setting`, `HRI_Type_Encountered`, `Sample_Type`, gender, age, and
`WEIRD_Sample` are all external-validity/sample-composition indicators and
are summarized together in the dedicated section below, not here.

## Helper function

```{r helper}
summarize_binary <- function(df, col, dimension, label, denom_filter = NULL) {
  x <- df[[col]]
  if (!is.null(denom_filter)) {
    x <- x[denom_filter]
  }
  n_applicable <- sum(x %in% c("Yes", "No"))
  n_yes <- sum(x == "Yes")
  pct <- if (n_applicable > 0) round(100 * n_yes / n_applicable, 1) else NA
  data.frame(
    Framework_Dimension = dimension,
    Indicator = label,
    N_Applicable = n_applicable,
    N_Yes = n_yes,
    Percent_Yes = pct,
    stringsAsFactors = FALSE
  )
}

has_subjective_dv <- df$Has_Subjective_DV == "Yes"
```

## Theoretical grounding

```{r theoretical-grounding}
tg <- rbind(
  summarize_binary(df, "Has_Hypothesis", "Theoretical grounding", "Paper states a hypothesis"),
  summarize_binary(df, "Theory_Driven_Hypothesis", "Theoretical grounding", "Hypothesis is theory-driven (of papers with a hypothesis)"),
  summarize_binary(df, "HARKing_Risk", "Theoretical grounding", "HARKing risk flagged (of papers with a hypothesis)")
)
knitr::kable(tg, row.names = FALSE)
```

## Internal validity

```{r internal-validity}
iv <- summarize_binary(df, "Random_Assignment_or_Order_Control_Reported",
                        "Internal validity",
                        "Random assignment / order control reported")
knitr::kable(iv, row.names = FALSE)
```

## Instrumentation validity

Denominator restricted to the `r sum(has_subjective_dv)` papers that include
at least one subjective (self-report) dependent variable.

```{r instrumentation-validity}
inst <- rbind(
  summarize_binary(df, "Has_Subjective_DV", "Instrumentation validity", "Paper includes >=1 subjective DV"),
  summarize_binary(df, "MultiItem_Measure_Used", "Instrumentation validity", "Multi-item measure used (of papers with a subjective DV)", has_subjective_dv),
  summarize_binary(df, "Existing_Measure_Used", "Instrumentation validity", "Existing/adapted measure used (of papers with a subjective DV)", has_subjective_dv),
  summarize_binary(df, "Reliability_Reported", "Instrumentation validity", "Reliability reported (of papers with a subjective DV)", has_subjective_dv),
  summarize_binary(df, "Construct_Validity_Reported", "Instrumentation validity", "Construct validity reported (of papers with a subjective DV)", has_subjective_dv)
)
knitr::kable(inst, row.names = FALSE)
```

## Manipulation validity

```{r manipulation-validity}
mv <- summarize_binary(df, "Manipulation_Check_Reported", "Manipulation validity",
                        "At least one manipulation check reported")
knitr::kable(mv, row.names = FALSE)
```

## External validity: sample and study characteristics

`Experiment_Setting`, `HRI_Type_Encountered`, and `Sample_Type` are summarized
here **by paper** (each paper counted exactly once, in exactly one category),
not with the "at least one experiment = Yes" rule used for the binary
indicators above. A paper whose multiple experiments used different
settings/types/populations is shown in its own explicit "Multiple" category,
and unreported values get their own "Not reported" category, so that every
column sums to its group's N. This matches the convention already used in
the Results-section overview paragraph.

```{r ev-helpers}
is_tech <- df$Technical_System_Paper == "Yes"

summarize_category <- function(x, cat_levels, is_tech) {
  x <- factor(x, levels = cat_levels)
  make_col <- function(idx) {
    t <- table(x[idx])
    n <- sum(idx)
    sprintf("%d (%.1f%%)", as.integer(t), 100 * as.integer(t) / n)
  }
  data.frame(
    Category = cat_levels,
    All = make_col(rep(TRUE, nrow(df))),
    Technical = make_col(is_tech),
    NonTechnical = make_col(!is_tech),
    stringsAsFactors = FALSE
  )
}
```

### Study setting

```{r ev-setting}
classify_setting <- function(x) {
  x <- ifelse(is.na(x) | x == "", "Not reported", x)
  ifelse(grepl("; ", x), "Multiple (differs across experiments)", x)
}
setting_levels <- c("Lab experiment", "Field experiment", "Online experiment",
                     "Not reported", "Multiple (differs across experiments)")
setting_tbl <- summarize_category(classify_setting(df$Experiment_Setting), setting_levels, is_tech)
knitr::kable(setting_tbl, row.names = FALSE,
             caption = paste0("Study setting, by paper (All n=", nrow(df),
                               ", Technical n=", sum(is_tech),
                               ", Non-technical n=", sum(!is_tech), ")"))
```

### HRI type encountered

```{r ev-hritype}
classify_hritype <- function(x) {
  x <- ifelse(is.na(x) | x == "", "Other or information not provided", x)
  ifelse(grepl("; ", x), "Multiple (differs across experiments)", x)
}
hritype_levels <- c("Actual Physical Robot Interaction (Real robot)",
                     "Video/Description of Robot shown to participants",
                     "Actual Virtual Robot Interaction (AR/VR)",
                     "Other or information not provided",
                     "Multiple (differs across experiments)")
hritype_tbl <- summarize_category(classify_hritype(df$HRI_Type_Encountered), hritype_levels, is_tech)
knitr::kable(hritype_tbl, row.names = FALSE, caption = "HRI type encountered, by paper")
```

### Sample / population type

`Sample_Type` cells can combine more than one population within a single
experiment (comma-separated, e.g. "Students, Employees") or differ across a
paper's multiple experiments (semicolon-separated). These are kept as
distinct categories below rather than collapsed together, since they reflect
different things (a genuinely mixed sample vs. inconsistent
sampling/reporting across experiments).

```{r ev-sampletype}
classify_sampletype <- function(x) {
  x <- ifelse(is.na(x) | x == "", "Information not provided", x)
  out <- ifelse(grepl("; ", x), "Multiple (differs across experiments)", x)
  out <- ifelse(out != "Multiple (differs across experiments)" & grepl(",", out),
                "Mixed sample (multiple population types)", out)
  out
}
sampletype_levels <- c("General Population", "Students", "Employees", "Retirement Community",
                        "Mixed sample (multiple population types)",
                        "Multiple (differs across experiments)", "Information not provided")
st_tbl <- summarize_category(classify_sampletype(df$Sample_Type), sampletype_levels, is_tech)
knitr::kable(st_tbl, row.names = FALSE, caption = "Participant population/sample type, by paper")
```

### Gender composition

```{r ev-gender}
female <- suppressWarnings(as.numeric(df$Female_N))
male   <- suppressWarnings(as.numeric(df$Male_N))
other  <- suppressWarnings(as.numeric(df$Other_Gender_N)); other[is.na(other)] <- 0
gender_total <- ifelse(is.na(female), 0, female) + ifelse(is.na(male), 0, male) + other
gender_reported <- (!is.na(female) | !is.na(male)) & gender_total > 0
pct_female <- ifelse(gender_reported, 100 * ifelse(is.na(female), 0, female) / gender_total, NA)
pct_male   <- ifelse(gender_reported, 100 * ifelse(is.na(male), 0, male) / gender_total, NA)
pct_other  <- ifelse(gender_reported, 100 * other / gender_total, NA)

gender_summary <- function(idx) {
  n <- sum(idx)
  rep_n <- sum(gender_reported[idx], na.rm = TRUE)
  data.frame(
    N = n,
    Gender_Reported = sprintf("%d (%.1f%%)", rep_n, 100 * rep_n / n),
    Mean_Pct_Female = round(mean(pct_female[idx], na.rm = TRUE), 1),
    Median_Pct_Female = round(median(pct_female[idx], na.rm = TRUE), 1),
    Mean_Pct_Male = round(mean(pct_male[idx], na.rm = TRUE), 1),
    Median_Pct_Male = round(median(pct_male[idx], na.rm = TRUE), 1),
    Mean_Pct_Other = round(mean(pct_other[idx], na.rm = TRUE), 1),
    Median_Pct_Other = round(median(pct_other[idx], na.rm = TRUE), 1)
  )
}
gender_tbl <- rbind(
  cbind(Group = "All", gender_summary(rep(TRUE, nrow(df)))),
  cbind(Group = "Technical", gender_summary(is_tech)),
  cbind(Group = "Non-technical", gender_summary(!is_tech))
)
knitr::kable(gender_tbl, row.names = FALSE,
             caption = "Gender reporting, and mean/median % female, % male, and % other-gender among papers that reported a gender breakdown")
```

### Participant age

```{r ev-age}
mean_age <- suppressWarnings(as.numeric(df$Mean_Age))
age_reported <- !is.na(mean_age)

age_summary <- function(idx) {
  n <- sum(idx)
  rep_n <- sum(age_reported[idx])
  vals <- mean_age[idx & age_reported]
  data.frame(
    N = n,
    Age_Reported = sprintf("%d (%.1f%%)", rep_n, 100 * rep_n / n),
    Mean = round(mean(vals), 1),
    Median = round(median(vals), 1),
    Min = round(min(vals), 1),
    Max = round(max(vals), 1)
  )
}
age_tbl <- rbind(
  cbind(Group = "All", age_summary(rep(TRUE, nrow(df)))),
  cbind(Group = "Technical", age_summary(is_tech)),
  cbind(Group = "Non-technical", age_summary(!is_tech))
)
knitr::kable(age_tbl, row.names = FALSE,
             caption = "Reported participant age (paper-level Mean_Age across the papers that reported it)")
```

### Generalizability limitation discussed (of WEIRD papers)

Among papers classified as WEIRD, whether the paper discusses the resulting
limitation to generalizability. `WEIRD_Limitation_Acknowledged` values of
"Yes" and "Partially" are both treated as "Yes" here; "No" stays "No".

```{r ev-weird}
is_weird <- df$WEIRD_Sample == "Yes"
classify_limitation_discussed <- function(x) ifelse(x %in% c("Yes", "Partially"), "Yes", "No")

limitation_summary <- function(idx) {
  sub_idx <- idx & is_weird
  n <- sum(sub_idx)
  x <- factor(classify_limitation_discussed(df$WEIRD_Limitation_Acknowledged[sub_idx]), levels = c("Yes", "No"))
  t <- table(x)
  data.frame(
    N_WEIRD = n,
    Yes = sprintf("%d (%.1f%%)", t["Yes"], 100 * t["Yes"] / n),
    No = sprintf("%d (%.1f%%)", t["No"], 100 * t["No"] / n)
  )
}
limitation_tbl <- rbind(
  cbind(Group = "All", limitation_summary(rep(TRUE, nrow(df)))),
  cbind(Group = "Technical", limitation_summary(is_tech)),
  cbind(Group = "Non-technical", limitation_summary(!is_tech))
)
knitr::kable(limitation_tbl, row.names = FALSE,
             caption = "Generalizability limitation discussed, among papers classified as WEIRD")
```

### Geography (participant country)

`Participant_Country` was fixed on 2026-09-16: the build script had been
joining two different master-codebook columns together (a Yes/No "was a
country reported" flag, and the actual free-text country field), producing
nonsense values like `"Japan; Yes"` or a bare `"No"`. It now holds only the
free-text country/description, with unreported cells as `"Not reported"`
(see Notes item 15 in the Data_Dictionary workbook).

Standard adopted (Notes item 16): `Participant_Country` = "Reported" only
when the paper explicitly names a specific country/countries. An inferred
country (from language, materials, or author affiliation) or a confirmed-
multinational-but-unenumerated sample both count as "Not reported", since
neither is a direct report of a specific country. 4 papers were re-read
against the source PDFs under this standard and are coded "Not reported":
Paper_ID 19 and 76 (a country is strongly implied but never stated), and
Paper_ID 25 and 33 (the paper confirms a multinational sample but never
lists the countries). `"NA"`, `"NR"`, and a bare `"No"` are also treated as
"Not reported" (spelling variants used by different coders/experiments in
the original sheet).

```{r ev-geography}
not_reported_tokens <- c("not reported", "na", "n/a", "nr", "no")
classify_country <- function(x) {
  x <- ifelse(is.na(x) | x == "", "Not reported", x)
  ifelse(tolower(trimws(x)) %in% not_reported_tokens, "Not reported", "Reported")
}
country_levels <- c("Reported", "Not reported")
country_tbl <- summarize_category(classify_country(df$Participant_Country), country_levels, is_tech)
knitr::kable(country_tbl, row.names = FALSE,
             caption = "Whether a specific participant country was explicitly reported, by paper")

knitr::kable(as.data.frame(table(Participant_Country = df$Participant_Country)),
             caption = "Participant_Country - raw values, for reference")
```

```{r ev-summary-export}
external_validity_summary <- rbind(
  cbind(Dimension = "Study setting", setting_tbl),
  cbind(Dimension = "HRI type encountered", hritype_tbl),
  cbind(Dimension = "Sample/population type", st_tbl),
  cbind(Dimension = "Participant country reported", country_tbl)
)
write.csv(external_validity_summary, "results_summary_external_validity.csv", row.names = FALSE)
```

## Statistical conclusion validity

```{r statistical-conclusion-validity}
scv <- rbind(
  summarize_binary(df, "Sample_Size_Justified", "Statistical conclusion validity", "Sample size justified (power analysis or other rationale)"),
  summarize_binary(df, "Assumptions_Checked", "Statistical conclusion validity", "Statistical test assumptions checked/reported"),
  summarize_binary(df, "Analysis_Issue_Identified", "Statistical conclusion validity", "An analysis issue was identified (e.g., uncorrected multiple comparisons)"),
  summarize_binary(df, "Effect_Size_Reported", "Statistical conclusion validity", "Effect size reported")
)
knitr::kable(scv, row.names = FALSE)
```

## Research transparency

```{r research-transparency}
rt <- rbind(
  summarize_binary(df, "Preregistration", "Research transparency", "Preregistration reported"),
  summarize_binary(df, "IRB_Reported", "Research transparency", "IRB/ethics approval reported")
)
knitr::kable(rt, row.names = FALSE)
```

## Combined summary table

```{r combined-summary}
results_summary <- rbind(tg, iv, inst, mv, scv, rt)
knitr::kable(results_summary, row.names = FALSE,
             caption = "All framework-dimension indicators (Table for Results section)")
write.csv(results_summary, "results_summary_by_dimension.csv", row.names = FALSE)
```

## Technical/Systems papers vs. other papers

Papers are split by `Technical_System_Paper` (Yes = the paper proposed a new
system/design/algorithm and the user study mainly evaluates it; No =
User Studies, Design, or Theory and Methods papers).

```{r tech-vs-nontech}
is_tech <- df$Technical_System_Paper == "Yes"

summarize_by_group <- function(df, group, col, dimension, label, denom_filter = NULL) {
  base_filter <- if (is.null(denom_filter)) rep(TRUE, nrow(df)) else denom_filter
  yes_group <- summarize_binary(df[group & base_filter, ], col, dimension, label)
  no_group  <- summarize_binary(df[!group & base_filter, ], col, dimension, label)
  data.frame(
    Framework_Dimension = dimension,
    Indicator = label,
    Technical_N = yes_group$N_Applicable, Technical_Pct_Yes = yes_group$Percent_Yes,
    NonTechnical_N = no_group$N_Applicable, NonTechnical_Pct_Yes = no_group$Percent_Yes,
    stringsAsFactors = FALSE
  )
}

has_subj_df <- df[has_subjective_dv, ]
is_tech_subj <- is_tech[has_subjective_dv]

comparison <- rbind(
  summarize_by_group(df, is_tech, "Has_Hypothesis", "Theoretical grounding", "Paper states a hypothesis"),
  summarize_by_group(df, is_tech, "Random_Assignment_or_Order_Control_Reported", "Internal validity", "Random assignment / order control reported"),
  summarize_by_group(has_subj_df, is_tech_subj, "MultiItem_Measure_Used", "Instrumentation validity", "Multi-item measure used"),
  summarize_by_group(has_subj_df, is_tech_subj, "Reliability_Reported", "Instrumentation validity", "Reliability reported"),
  summarize_by_group(has_subj_df, is_tech_subj, "Construct_Validity_Reported", "Instrumentation validity", "Construct validity reported"),
  summarize_by_group(df, is_tech, "Manipulation_Check_Reported", "Manipulation validity", "At least one manipulation check reported"),
  summarize_by_group(df, is_tech, "Sample_Size_Justified", "Statistical conclusion validity", "Sample size justified"),
  summarize_by_group(df, is_tech, "Effect_Size_Reported", "Statistical conclusion validity", "Effect size reported"),
  summarize_by_group(df, is_tech, "Preregistration", "Research transparency", "Preregistration reported")
)
knitr::kable(comparison, row.names = FALSE,
             caption = paste0("Technical/Systems papers (n = ", sum(is_tech),
                               ") vs. other papers (n = ", sum(!is_tech), ")"))
write.csv(comparison, "results_summary_technical_vs_nontechnical.csv", row.names = FALSE)
```
